The impact of cosmic variance on simulating weak lensing surveys

### 2014-10-24 Writeup/paper_v3.tex
Figures modified cosmetically

Once we volume-limit our sample (explained in Sec.~\ref{sub:volumelimiting}), we again compare the histogram to the models.
If there is no qualitative change in the overdensities, we stick with the preliminary decision. If the overdensity values
flip in sign or become too small, then we classify the redshift bin as `neutral'.

\arun{New:}The naive Poisson uncertainty of the counts
in each bin is much lesser compared the difference between the actual number of galaxies present
and the number predicted by the models. For this reason, errorbars have been completely ignored.

@@

%Alternatively, we can impose cuts on stellar mass of the galaxies instead of imposing on the luminosity. 
%If we restrict to galaxies with masses $M$ such that $\log(M/M_\odot) \ge 10.15$, then we have at least {\bf 95.0\% } completeness.
Finally, we can circumvent the problem of redshift evolution of the
luminosity by imposing cuts on stellar mass instead. In Fig.~\ref{fig:smf}, we show the stellar mass function (SMF) of our sample for various F814W flux limits.
\cite{Tomczak_SMF} report the SMFs for the ZFOURGE survey, which
includes COSMOS. They calculated stellar masses using the procedure
and software described by \cite{2009ApJ...700..221K}, using a set of models
with exponentially declining star formation history \citep{2003MNRAS.344.1000B} assuming a
Chabrier IMF \citep{ChabrierIMF}.  (\rachel{What do you mean by a single stellar population?} \arun{I had lifted that term from the paper since I thought it was important to highlight that. 
They consider those stars that have metallicity of that of the sun
formed at redshift $z=5$.} \rachel{No, look more carefully: they used
that only to estimate completeness.  Their description of the stellar
mss estimation process says they used a set of BC03
models.  This would never work for a single population!  I left this
conversation here to correct this misimpression, but since I
fixed the text already, please remove this conversation once you've
read it - there are no further action items.}) 
We plot their SMF for \emph{all} galaxies in Fig.~\ref{fig:smf} for
reference. Their SMF is higher than ours since they reach a $K_s$-band
$5\sigma$ depth of 24.9. 
As for $M_I$, we compare the stellar mass function in the $m_\text{F814W}\le23.5$ sample with that in the $m_\text{F814W}\le25.2$ sample. 
The sample with $\log(M/M_\odot) > 10.15$ is $\sim 95$ per cent
complete in the redshift bin $0.75 \le z< 0.85$ and has 10~341
galaxies in total across all redshifts.
Thus, we construct a third volume-limited sample \s$3$ by imposing the
stellar mass cut mentioned above.

@@

The functional forms for the (flux-limited) redshift distribution that
we used in Sec.~\ref{sub:overdensities} are not well-motivated for a
volume-limited sample. If we fit them to the 
redshift distribution of the volume-limited sample ($M_I < -22$
\rachel{without or with evolution correction?}), then 
due to the absence of an exponential tail in the histogram, the
parameters that set the scale for the redshift ($z_1$ and $z_2$)
become very large.  As a result, both  $p_1(z)$ and $p_2(z)$ defined
in Eqs.~\eqref{E:pz1} and~\eqref{E:pz2} essentially become the same power law.
The best-fitting exponent is remarkably close to $2$ ($a=2.78\pm
0.28$, giving an exponent $1.78\pm 0.28$), suggesting that the
comoving number density of galaxies is constant with redshift as we
would expect for a volume-limited sample in a redshift range for which
evolution is negligible. 
Fig.~\ref{fig:redshift_vollimited} shows that when we do
this, the values of $\delta_{g,\text{1D}}$ for the $z=0.40-0.55$ bin
increase and are within the $[-0.1, 0.1]$ range that we have defined
as neutral. This is the reason that in Sec.~\ref{sub:overdensities} we classified them as neutral as opposed to underdense.
We will see in Sec.~\ref{S:results} that they are more similar to overdense regions as opposed to underdense regions.
The other redshift slices seem to exhibit a consistent behavior in
Fig.~\ref{fig:redshift_vollimited} as in Fig.~\ref{fig:redshift_fluxlimited}.

@@

% \rachel{As we discussed, you should get rid of the early parts of this
%   regarding second moments, and replace it with a brief description of
%   what shape estimates we use from GalSim (re-Gaussianization).}
% 
% {\bf Should explain in brief Claire's work.} \\
\begin{figure*}
 \centering
 \includegraphics[width=2.2\columnwidth]{sample_histograms}
 \caption{The distributions of the morphological parameters of interest, namely the distortion, \sersic index and \btt are presented. 
          The shape of the distribution depends on the way the volume-limiting process is carried out. \s$0$ refers to the flux-limited ($m_\text{F814}\le23.5$) sample
          and \s$1$, \s$2$ and \s$3$ refer to the volume-limited
          samples, as discussed in
          Sec.~\ref{sub:volumelimiting}. \rachel{I am confused by the
            top line: first, distortion is $(1-q^2)/(1+q^2)$, not
            $(1-q)/(1+q)$.  Second, I am not sure which shapes you are
          plotting - from Claire or from re-Gaussianization.  It seems
        redundant to have two plots of the same thing, so I hope that
        one is from Claire and one from re-Gaussianization, but these
        needs to be indicated clearly.  Finally, we have the option of
      deliberately plotting different things (like $q$ from Claire
      vs. distortion from re-Gaussianization) or explicitly choosing
      to plot the same thing (like total distortion from Claire
      vs. total distortion from re-Gaussianization).  I'm not sure
      which is best; the latter would allow a comparison between the
      two types of shape estimates, which could be interesting, so I'm
      inclined to say you should do that.}}
  \label{fig:sample_histograms}
\end{figure*}

@@

\subsubsection{RMS ellipticites}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip1_noevolution.eps} \\
 \includegraphics[width=1.0\columnwidth]{rms_ellip1_Bbandevolution.eps} \\
% \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \caption{RMS ellipticities as a function of redshift. The horizontal errorbars indicate
          the width of the redshift bin, while the vertical ones
          are $1\sigma$ errorbars obtained by
          bootstrapping. The colors and shapes of the points indicate their
          environmental classification, as shown in the legend. Points with open centers and thick errorbars correspond to the stellar-mass selected sample \s$3$
          and points with filled centers and thin errorbars correspond to the luminosity-selected samples \s$1$ and \s$2$.             
          }
 \label{fig:rms_ellip}

@@

\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{redshift_fluxlimited_wide}
 \includegraphics[width=1.0\columnwidth]{redshift_vollimited_wide}
 \caption{Plot of $\delta_{g,\text{1D}} =  N/N_{\text{mod}}-1$ with each functional form as the model
           for each of our new wide redshift bins discussed in Sec.~\ref{sec:mitigation}, for flux-limited (top) and
           volume-limited (bottom) samples. \rachel{Remove $p_2(z)$
             in bottom panel since it's hidden anyway so seeing it in the legend is
             just confusing.}}
 \label{fig:redshift_wide}          
\end{figure}

@@

Using the redshift distributions estimated earlier in this work (cf. Figs.~\ref{fig:redshift_fluxlimited}
and~\ref{fig:redshift_vollimited}), we obtain the overdensity estimates for
the wider bins, $\delta_{g,\text{1D}}$.
%that are shown in Figs.~\ref{fig:redshift_fluxlimited_wide} and ~\ref{fig:redshift_vollimited_wide}. 
Although the $\left[ 0.7 - 0.85\right]$ bin seems to be overdense in
the top panel of Fig.~\ref{fig:redshift_wide}, it appears to be
environmentally neutral when we volume-limit the sample (\rachel{how?
  Should say here and in caption whether this is \s1, \s2, or \s3}).
Since the latter is what we use to study the galaxy morphology, we
classify $\left[ 0.7 - 0.85\right]$ bin as `neutral'.
What is surprising is the fact that the lowest and middle redshift
slices still qualify as substantial overdensities and underdensities
despite our use of $\Delta z=0.2$.

>>>>

Once we volume-limit our sample (explained in Sec.~\ref{sub:volumelimiting}), we again compare the histogram to the models.
If there is no qualitative change in the overdensities, we stick with the preliminary decision. If the overdensity values
flip in sign or become too small, then we classify the redshift bin as `neutral'.
%\arun{New:}
The naive Poisson uncertainty of the counts
in each bin is much lesser compared the difference between the actual number of galaxies present
and the number predicted by the models. For this reason, errorbars have been completely ignored.

@@

%Alternatively, we can impose cuts on stellar mass of the galaxies instead of imposing on the luminosity. 
%If we restrict to galaxies with masses $M$ such that $\log(M/M_\odot) \ge 10.15$, then we have at least {\bf 95.0\% } completeness.
Finally, we can circumvent the problem of redshift evolution of the
luminosity by imposing cuts on stellar mass instead. In Fig.~\ref{fig:smf}, we show the stellar mass function (SMF) of our sample for various F814W flux limits.
\cite{Tomczak_SMF} report the SMFs for the ZFOURGE survey, which
includes COSMOS. They calculated stellar masses using the procedure
and software described by \cite{2009ApJ...700..221K}, using a set of models
with exponentially declining star formation history \citep{2003MNRAS.344.1000B} assuming a
Chabrier IMF \citep{ChabrierIMF}.  (\rachel{What do you mean by a single stellar population?} \arun{I had lifted that term from the paper since I thought it was important to highlight that. 
They consider those stars that have metallicity of that of the sun
formed at redshift $z=5$.} 
% \rachel{No, look more carefully: they used
% that only to estimate completeness.  Their description of the stellar
% mss estimation process says they used a set of BC03
% models.  This would never work for a single population!  I left this
% conversation here to correct this misimpression, but since I
% fixed the text already, please remove this conversation once you've
% read it - there are no further action items.}) 
We plot their SMF for \emph{all} galaxies in Fig.~\ref{fig:smf} for
reference. Their SMF is higher than ours since they reach a $K_s$-band
$5\sigma$ depth of 24.9. 
As for $M_I$, we compare the stellar mass function in the $m_\text{F814W}\le23.5$ sample with that in the $m_\text{F814W}\le25.2$ sample. 
The sample with $\log(M/M_\odot) > 10.15$ is $\sim 95$ per cent
complete in the redshift bin $0.75 \le z< 0.85$ and has 10~341
galaxies in total across all redshifts.
Thus, we construct a third volume-limited sample \s$3$ by imposing the
stellar mass cut mentioned above.

@@

The functional forms for the (flux-limited) redshift distribution that
we used in Sec.~\ref{sub:overdensities} are not well-motivated for a
volume-limited sample. If we fit them to the 
redshift distribution of the volume-limited sample ($M_I < -22$) that doesn't take into account the evolution of the luminosity function,
%\rachel{without or with evolution correction?}\arun{without} 
then due to the absence of an exponential tail in the histogram, the
parameters that set the scale for the redshift ($z_1$ and $z_2$)
become very large.  As a result, both  $p_1(z)$ and $p_2(z)$ defined
in Eqs.~\eqref{E:pz1} and~\eqref{E:pz2} essentially become the same power law.
The best-fitting exponent is remarkably close to $2$ ($a=2.78\pm
0.28$, giving an exponent $1.78\pm 0.28$), suggesting that the
comoving number density of galaxies is constant with redshift as we
would expect for a volume-limited sample in a redshift range for which
evolution is negligible. 
Fig.~\ref{fig:redshift_vollimited} shows that when we do
this, the values of $\delta_{g,\text{1D}}$ for the $z=0.40-0.55$ bin
increase and are within the $[-0.1, 0.1]$ range that we have defined
as neutral. This is the reason that in Sec.~\ref{sub:overdensities} we classified them as neutral as opposed to underdense.
We will see in Sec.~\ref{S:results} that they are more similar to overdense regions as opposed to underdense regions.
The other redshift slices seem to exhibit a consistent behavior in
Fig.~\ref{fig:redshift_vollimited} as in Fig.~\ref{fig:redshift_fluxlimited}.

@@

% \rachel{As we discussed, you should get rid of the early parts of this
%   regarding second moments, and replace it with a brief description of
%   what shape estimates we use from GalSim (re-Gaussianization).}
% 
% {\bf Should explain in brief Claire's work.} \\
\begin{figure*}
 \centering
 \includegraphics[width=2.2\columnwidth]{sample_histograms}
 \caption{The distributions of the morphological parameters of interest, namely the distortion, \sersic index and \btt are presented. 
          The shape of the distribution depends on the way the volume-limiting process is carried out. \s$0$ refers to the flux-limited ($m_\text{F814}\le23.5$) sample
          and \s$1$, \s$2$ and \s$3$ refer to the volume-limited
          samples, as discussed in
          Sec.~\ref{sub:volumelimiting}. 
%           \rachel{I am confused by the
%             top line: first, distortion is $(1-q^2)/(1+q^2)$, not
%             $(1-q)/(1+q)$.  Second, I am not sure which shapes you are
%           plotting - from Claire or from re-Gaussianization.  It seems
%         redundant to have two plots of the same thing, so I hope that
%         one is from Claire and one from re-Gaussianization, but these
%         needs to be indicated clearly.  Finally, we have the option of
%       deliberately plotting different things (like $q$ from Claire
%       vs. distortion from re-Gaussianization) or explicitly choosing
%       to plot the same thing (like total distortion from Claire
%       vs. total distortion from re-Gaussianization).  I'm not sure
%       which is best; the latter would allow a comparison between the
%       two types of shape estimates, which could be interesting, so I'm
%       inclined to say you should do that.}} \arun{Done.}
  }
  \label{fig:sample_histograms}
\end{figure*}

@@

\subsubsection{RMS ellipticites}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \includegraphics[width=1.0\columnwidth]{rms_ellip2_Bbandevolution.eps} \\
% \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \caption{RMS ellipticities as a function of redshift. The horizontal errorbars indicate
          the width of the redshift bin, while the vertical ones
          are $1\sigma$ errorbars obtained by
          bootstrapping. The colors and shapes of the points indicate their
          environmental classification, as shown in the legend. Points with open centers and thick errorbars correspond to the stellar-mass selected sample \s$3$
          and points with filled centers and thin errorbars correspond to the luminosity-selected samples \s$1$ and \s$2$.             
          }
 \label{fig:rms_ellip}

@@

\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{redshift_fluxlimited_wide}
 \includegraphics[width=1.0\columnwidth]{redshift_vollimited_wide}
 \caption{Plot of $\delta_{g,\text{1D}} =  N/N_{\text{mod}}-1$ with each functional form as the model
           for each of our new wide redshift bins discussed in Sec.~\ref{sec:mitigation}, for flux-limited (top) and
           volume-limited (bottom) samples.
           %\rachel{Remove $p_2(z)$             in bottom panel since it's hidden anyway so seeing it in the legend is             just confusing.} \arun{Instead, thickened the line.}
          }
 \label{fig:redshift_wide}          
\end{figure}

@@

Using the redshift distributions estimated earlier in this work (cf. Figs.~\ref{fig:redshift_fluxlimited}
and~\ref{fig:redshift_vollimited}), we obtain the overdensity estimates for
the wider bins, $\delta_{g,\text{1D}}$.
%that are shown in Figs.~\ref{fig:redshift_fluxlimited_wide} and ~\ref{fig:redshift_vollimited_wide}. 
Although the $\left[ 0.7 - 0.85\right]$ bin seems to be overdense in
the top panel of Fig.~\ref{fig:redshift_wide}, it appears to be
environmentally neutral when we volume-limit the sample to be complete upto $z=1$.
%(\rachel{how?  Should say here and in caption whether this is \s1, \s2, or \s3}).
Since the latter is what we use to study the galaxy morphology, we
classify $\left[ 0.7 - 0.85\right]$ bin as `neutral'.
What is surprising is the fact that the lowest and middle redshift
slices still qualify as substantial overdensities and underdensities
despite our use of $\Delta z=0.2$.
